---
title: "ISSS608-VAA_demo"
format: html
---

Welcome to ISSS608 Visual Analytics and Applications. In this website, you will find my coursework prepared for this course.

## Importing Exam Data

```{r}
library(readr)

exam_data <- read_csv("data/Exam_data.csv")

exam_data

## English versus Maths scores

```{r}
library(ggplot2)
library(ggrepel)

ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5) +
  geom_label_repel(aes(label = ID),
                   size = 2.5) +
  labs(title = "English scores versus Maths scores for Primary 3",
       x = "MATHS",
       y = "ENGLISH")
```
## English versus Maths scores

```{r}
library(ggplot2)
library(ggrepel)

ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = "lm", 
              linewidth = 0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
## Distribution of Maths scores

```{r}
library(ggplot2)

ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```
## Distribution of Maths scores with Economist theme

```{r}
library(ggplot2)
library(ggthemes)

ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
## Distribution of Maths scores with Ipsum theme

```{r}
library(ggplot2)
library(hrbrthemes)

ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
## Distribution of Maths scores with Ipsum theme

```{r}
library(ggplot2)
library(hrbrthemes)

ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```
## Patchwork Layouts

```{r}
library(ggplot2)
library(patchwork)

p1 <- ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") + 
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of Maths scores")

p2 <- ggplot(data = exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of English scores")

p3 <- ggplot(data = exam_data, 
             aes(x = MATHS, 
                 y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = "lm", 
              linewidth = 0.5) +  
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

(p1 / p2) | p3

((p1 / p2) | p3) +
  plot_annotation(tag_levels = 'I')

p3 + inset_element(p2,
                   left = 0.02,
                   bottom = 0.7,
                   right = 0.5,
                   top = 1)
```