---
title: "Take-home Exercise 1: Visual Analytics of a Motor Insurance Portfolio"
date: "May 23, 2026"
editor: visual
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
---

# 1. Overview

## 1.1 Background

Motor insurers manage portfolios where the overall result can look acceptable while risk is concentrated in a small number of policies, coverages, vehicle profiles, or claim events. Premium adequacy is therefore not only about the average loss ratio. It also depends on whether the insurer can identify segments with higher claim frequency, higher claim severity, and more volatile incurred losses.

This study applies exploratory visual analytics to a real-world motor insurance portfolio. The purpose is to move beyond a dashboard-style summary and use fit-for-purpose visualisation methods to understand portfolio composition, claims behaviour, premium adequacy, and risk concentration.

## 1.2 The Data

Two files are used in this study:

- `Dataset of motor insurance portfolio.csv`: policy-year level observations of motor insurance contracts, premiums, exposure, claim counts, and incurred costs.
- `Descriptive of variables.xlsx`: a variable dictionary describing the fields in the portfolio dataset.

The dataset contains one row per policy-year observation. The same `insured_id` can therefore appear across multiple calendar years.

## 1.3 The Task

Using the motor insurance portfolio dataset, this report aims to use appropriate exploratory data analysis methods and `ggplot2` visualisations to uncover:

- the distribution of policy, driver, vehicle, premium, claim, and incurred loss variables;
- the relationship between profitability or volatility and policy, underwriting, vehicle, and geographic characteristics; and
- practical portfolio insights that can support underwriting, pricing, and claims monitoring.

# 2. Loading Packages

The report uses a compact set of R packages, included : package loading, data preparation, univariate analysis, exploratory data analysis, and a concluding summary.

```{r}
library(tidyverse)
library(readxl)
library(knitr)
library(patchwork)
library(ggthemes)
library(scales)
library(ggridges)
library(ggdist)
library(plotly)
library(ggiraph)
library(DT)
```

```{r}
package_tbl <- tribble(
  ~Package, ~Purpose,
  "tidyverse", "Data import, wrangling, summarisation, and ggplot2 visualisation.",
  "readxl", "Importing the Excel variable dictionary.",
  "knitr", "Rendering compact summary tables.",
  "patchwork", "Combining several ggplot charts into one analytical view.",
  "ggthemes", "Additional clean plotting themes.",
  "scales", "Readable axis labels for proportions, counts, and monetary values.",
  "ggridges", "Comparing full distributions across policy segments.",
  "ggdist", "Additional distribution geoms when needed.",
  "plotly", "Interactive charts with hover labels, zooming, and pan controls.",
  "ggiraph", "Interactive ggplot layers with tooltips and hover highlighting.",
  "DT", "Searchable, sortable, and filterable data tables."
)

kable(package_tbl)
```

```{r}
theme_set(theme_minimal(base_size = 11))

base_fill <- "#73b2c4"
accent_fill <- "#f27279"
dark_fill <- "#244c5a"
neutral_fill <- "#c7c8cc"
positive_fill <- "#5c9f6f"
warning_fill <- "#d8894c"

fmt_count <- label_number(big.mark = ",", accuracy = 1)
fmt_rate <- label_number(accuracy = 0.01)
fmt_pct <- label_percent(accuracy = 0.1)
fmt_money <- label_number(big.mark = ",", accuracy = 1)
```

# 3. Motor Insurance Portfolio Data

```{r}
data_dir <- file.path("data", "A detailed dataset of motor insurance policies wit")
portfolio_file <- file.path(data_dir, "Dataset of motor insurance portfolio.csv")
dictionary_file <- file.path(data_dir, "Descriptive of variables.xlsx")

motor_raw <- read_delim(
  portfolio_file,
  delim = ";",
  show_col_types = FALSE,
  progress = FALSE
)

var_dictionary <- read_excel(dictionary_file, sheet = "Cartera") %>%
  fill(Group, .direction = "down") %>%
  filter(!is.na(Variables)) %>%
  transmute(
    variable = Variables,
    description = Description,
    group = Group
  )
```

```{r}
data_profile_tbl <- tibble(
  Measure = c(
    "Rows",
    "Variables",
    "Distinct insured IDs",
    "Calendar years",
    "Exact duplicated rows"
  ),
  Value = c(
    fmt_count(nrow(motor_raw)),
    fmt_count(ncol(motor_raw)),
    fmt_count(n_distinct(motor_raw$insured_id)),
    paste(sort(unique(motor_raw$year)), collapse = ", "),
    fmt_count(sum(duplicated(motor_raw)))
  )
)

kable(data_profile_tbl)
```

The imported portfolio contains `r fmt_count(nrow(motor_raw))` policy-year records across `r fmt_count(ncol(motor_raw))` variables. There are `r fmt_count(n_distinct(motor_raw$insured_id))` distinct insured IDs, confirming that some policies appear in more than one year. No exact duplicated rows are found.

# 4. Data Wrangling

The following flowchart shows how the main variable groups are used and how the key analytical metrics are derived.

```{mermaid}
flowchart LR
    A("Motor insurance portfolio CSV") --> B["Policy-year records"]
    C("Variable dictionary XLSX") --> D["Variable meanings and category labels"]
    B --> E["Policy and underwriting profile"]
    B --> F["Driver and vehicle profile"]
    B --> G["Premium, exposure, claims, incurred"]
    E --> H["Segment analysis"]
    F --> H
    G --> I["Derived metrics"]
    I --> I1["Loss ratio = incurred / premium"]
    I --> I2["Claim frequency = claims / exposure"]
    I --> I3["Claim severity = incurred / claims"]
    I --> I4["Pure premium = incurred / exposure"]
    H --> J["Visual analytics and portfolio insights"]
    I --> J
```

## 4.1 Variable Selection

The full dataset is retained for reproducibility, but the analysis focuses on variables that support portfolio composition, profitability, and volatility analysis.

```{r}
selected_variables <- tribble(
  ~Category, ~Variable, ~Why_it_is_used,
  "Policy", "year", "Time trend of portfolio growth and performance.",
  "Policy", "policy_type", "Coverage structure is expected to influence premium, frequency, and incurred losses.",
  "Policy", "policy_status", "Cancelled policies may behave differently from active policies.",
  "Policy", "business_type", "Separates new business from existing portfolio business.",
  "Policy", "payment_frequency", "Billing frequency can be associated with customer and policy characteristics.",
  "Policy", "bonus_score", "Prior claims experience is a core underwriting signal.",
  "Driver and vehicle", "driver_age", "Driver age is commonly related to claim frequency and severity.",
  "Driver and vehicle", "vehicle_age", "Vehicle age can influence repair cost, claim behaviour, and coverage choice.",
  "Driver and vehicle", "fuel_type", "Fuel type may proxy vehicle use and cost patterns.",
  "Driver and vehicle", "vehicle_value", "Vehicle value helps interpret premium and claim severity.",
  "Driver and vehicle", "vehicle_brand", "Brand helps detect concentration in vehicle segments.",
  "Geography", "municipality_type, circulation_area", "Geographic and driving environment segments may carry different risk.",
  "Premiums", "total_premium", "Denominator for loss ratio and indicator of pricing level.",
  "Claims", "total_claims", "Core claim count variable for frequency analysis.",
  "Claims", "total_incurred", "Core loss cost variable for profitability and volatility analysis.",
  "Exposure", "total_exposure", "Normalises claim and loss experience by time on risk."
)

kable(selected_variables)
```

## 4.2 Handling Missing Values

Missingness is small and concentrated in vehicle-related fields. `fuel_type` is converted to an explicit `Unknown` category so that these policies remain visible in categorical charts. Numeric missing values are retained and excluded only from charts or summaries that require those variables.

```{r}
missing_tbl <- motor_raw %>%
  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_n") %>%
  filter(missing_n > 0) %>%
  mutate(
    missing_pct = missing_n / nrow(motor_raw),
    missing_n = fmt_count(missing_n),
    missing_pct = fmt_pct(missing_pct)
  )

kable(missing_tbl)
```

The missing values are limited: `fuel_type` has the largest missing count, while `vehicle_value`, `vehicle_age`, and `age_driving_licence` have very small missing shares. Because the missingness is low, the main portfolio-level results are not sensitive to these fields.

## 4.3 Recoding Categorical Variables

The categorical variables are recoded from short codes into readable labels using the descriptions from the Excel dictionary.

```{r}
safe_divide <- function(num, den) {
  if_else(!is.na(den) & den > 0, num / den, NA_real_)
}

motor_analysis <- motor_raw %>%
  mutate(
    policy_type = fct_recode(
      factor(policy_type),
      "Third-party" = "TP",
      "Third-party + glass" = "TPG",
      "Third-party + add-ons" = "CC",
      "Comprehensive with excess" = "COMP_E",
      "Comprehensive without excess" = "COMP_N"
    ),
    policy_type = factor(
      policy_type,
      levels = c(
        "Third-party",
        "Third-party + glass",
        "Third-party + add-ons",
        "Comprehensive with excess",
        "Comprehensive without excess"
      )
    ),
    policy_status = fct_recode(
      factor(policy_status),
      "Active" = "A",
      "Cancelled" = "C"
    ),
    business_type = fct_recode(
      factor(business_type),
      "New business" = "NB",
      "Portfolio" = "P"
    ),
    payment_frequency = fct_recode(
      factor(payment_frequency),
      "Annual" = "A",
      "Semiannual" = "S",
      "Quarterly" = "Q"
    ),
    payment_frequency = factor(
      payment_frequency,
      levels = c("Annual", "Semiannual", "Quarterly")
    ),
    bonus_score = fct_recode(
      factor(bonus_score),
      "Favourable history" = "G",
      "Neutral history" = "N",
      "Poor history" = "B"
    ),
    bonus_score = factor(
      bonus_score,
      levels = c("Poor history", "Neutral history", "Favourable history")
    ),
    fuel_type = fct_recode(
      factor(fuel_type),
      "Diesel" = "D",
      "Gasoline" = "G"
    ),
    fuel_type = fct_na_value_to_level(fuel_type, level = "Unknown"),
    municipality_type = fct_recode(
      factor(municipality_type),
      "Inland" = "I",
      "Coastal" = "C",
      "Islands" = "IS"
    ),
    circulation_area = fct_recode(
      factor(circulation_area),
      "Urban" = "U",
      "Rural" = "R"
    )
  )
```

## 4.4 Data Type Conversion

The `year` variable is treated as an ordered time variable for trend charts, while categorical variables are treated as factors for ordered plotting.

```{r}
motor_analysis <- motor_analysis %>%
  mutate(
    year = factor(year, levels = sort(unique(year))),
    year_num = as.integer(as.character(year))
  )
```

## 4.5 Variable Creation

Actuarial portfolio analysis requires normalised metrics. Claim count and incurred cost are divided by exposure, premium, or claim count depending on the question.

```{r}
motor_analysis <- motor_analysis %>%
  mutate(
    claim_indicator = total_claims > 0,
    loss_ratio = safe_divide(total_incurred, total_premium),
    claim_frequency = safe_divide(total_claims, total_exposure),
    claim_severity = safe_divide(total_incurred, total_claims),
    pure_premium = safe_divide(total_incurred, total_exposure),
    premium_per_exposure = safe_divide(total_premium, total_exposure),
    driver_age_band = cut(
      driver_age,
      breaks = c(17, 25, 35, 45, 55, 65, Inf),
      labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+")
    ),
    vehicle_age_band = cut(
      vehicle_age,
      breaks = c(-1, 5, 10, 20, 30, 40, Inf),
      labels = c("0-5", "6-10", "11-20", "21-30", "31-40", "41+")
    ),
    claim_count_band = case_when(
      total_claims == 0 ~ "0",
      total_claims == 1 ~ "1",
      total_claims == 2 ~ "2",
      total_claims >= 3 ~ "3+",
      TRUE ~ NA_character_
    ),
    claim_count_band = factor(claim_count_band, levels = c("0", "1", "2", "3+"))
  )
```

The key metrics used in the analysis are:

```{r}
metric_definitions <- tribble(
  ~Metric, ~Formula, ~Interpretation,
  "Loss ratio", "total_incurred / total_premium", "Profitability indicator. Higher values indicate that claims consume more premium.",
  "Claim frequency", "total_claims / total_exposure", "Claim counts normalised by time on risk.",
  "Claim severity", "total_incurred / total_claims", "Average incurred cost per claim, calculated only where claims exist.",
  "Pure premium", "total_incurred / total_exposure", "Expected loss cost per exposure year.",
  "Premium per exposure", "total_premium / total_exposure", "Annualised premium level."
)

kable(metric_definitions)
```

# 5. Univariate Analysis

## 5.1 Portfolio-Level Performance

```{r}
portfolio_metrics <- motor_analysis %>%
  summarise(
    records = n(),
    unique_policies = n_distinct(insured_id),
    exposure = sum(total_exposure, na.rm = TRUE),
    premium = sum(total_premium, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    pure_premium = incurred / exposure,
    avg_premium_per_exposure = premium / exposure,
    no_claim_share = mean(total_claims == 0, na.rm = TRUE),
    no_incurred_share = mean(total_incurred == 0, na.rm = TRUE)
  )

portfolio_metrics_display <- portfolio_metrics %>%
  transmute(
    `Policy-year records` = fmt_count(records),
    `Total exposure` = fmt_count(exposure),
    `Total premium` = fmt_money(premium),
    `Total claims` = fmt_count(claims),
    `Total incurred` = fmt_money(incurred),
    `Loss ratio` = fmt_pct(loss_ratio),
    `Claim frequency` = fmt_rate(claim_frequency),
    `Claim severity` = fmt_money(claim_severity),
    `No-claim share` = fmt_pct(no_claim_share)
  )

kable(portfolio_metrics_display)
```

At portfolio level, the loss ratio is `r fmt_pct(portfolio_metrics$loss_ratio)`, with `r fmt_rate(portfolio_metrics$claim_frequency)` claims per exposure year. The high no-claim share of `r fmt_pct(portfolio_metrics$no_claim_share)` indicates a strongly zero-inflated claim process: most policies do not claim, while a minority of claim records drive the loss result.

## 5.2 Portfolio Composition

```{r}
policy_mix <- motor_analysis %>%
  count(policy_type, name = "records") %>%
  mutate(
    share = records / sum(records),
    policy_type = fct_reorder(policy_type, records)
  )

ggplot(policy_mix, aes(x = records, y = policy_type, fill = policy_type)) +
  geom_col(width = 0.7) +
  geom_text(
    aes(label = fmt_pct(share)),
    hjust = -0.15,
    size = 3.2
  ) +
  scale_x_continuous(
    labels = fmt_count,
    expand = expansion(mult = c(0, 0.13))
  ) +
  scale_fill_manual(values = c(
    "Third-party" = "#b7b7b7",
    "Third-party + glass" = "#9ab7c6",
    "Third-party + add-ons" = "#73b2c4",
    "Comprehensive with excess" = "#f2a65a",
    "Comprehensive without excess" = "#f27279"
  )) +
  labs(
    title = "Policy type mix",
    subtitle = "Third-party with add-ons is the dominant coverage structure.",
    x = "Policy-year records",
    y = NULL
  ) +
  guides(fill = "none") +
  theme(
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold")
  )
```

```{r}
make_share_bar <- function(data, var, title) {
  data %>%
    count({{ var }}, name = "records") %>%
    mutate(
      share = records / sum(records),
      segment = fct_reorder(as.factor({{ var }}), records)
    ) %>%
    ggplot(aes(x = records, y = segment)) +
    geom_col(fill = base_fill, width = 0.7) +
    geom_text(aes(label = fmt_pct(share)), hjust = -0.1, size = 3) +
    scale_x_continuous(
      labels = fmt_count,
      expand = expansion(mult = c(0, 0.18))
    ) +
    labs(title = title, x = NULL, y = NULL) +
    theme(
      panel.grid.major.y = element_blank(),
      plot.title = element_text(face = "bold", size = 10)
    )
}

p_bonus <- make_share_bar(motor_analysis, bonus_score, "Bonus score")
p_payment <- make_share_bar(motor_analysis, payment_frequency, "Payment frequency")
p_fuel <- make_share_bar(motor_analysis, fuel_type, "Fuel type")
p_area <- make_share_bar(motor_analysis, circulation_area, "Circulation area")

(p_bonus | p_payment) / (p_fuel | p_area)
```

The composition charts show that the portfolio is concentrated in favourable bonus-score policies and annual payment policies. Diesel vehicles and urban circulation also account for substantial shares, making them important when interpreting portfolio-level results.

## 5.3 Numerical Variables

```{r}
numeric_plot_data <- motor_analysis %>%
  transmute(
    `Driver age` = driver_age,
    `Vehicle age` = vehicle_age,
    `Vehicle value` = vehicle_value,
    `Premium per exposure` = premium_per_exposure
  ) %>%
  pivot_longer(everything(), names_to = "metric", values_to = "value") %>%
  filter(!is.na(value), is.finite(value)) %>%
  group_by(metric) %>%
  mutate(
    p99 = quantile(value, 0.99, na.rm = TRUE),
    value_capped = pmin(value, p99)
  ) %>%
  ungroup()

ggplot(numeric_plot_data, aes(x = value_capped)) +
  geom_histogram(bins = 45, fill = base_fill, colour = "white", linewidth = 0.15) +
  facet_wrap(~ metric, scales = "free", ncol = 2) +
  scale_x_continuous(labels = fmt_money) +
  labs(
    title = "Distributions of key numerical variables",
    subtitle = "Values are capped at the 99th percentile only for display.",
    x = NULL,
    y = "Policy-year records"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )
```

The portfolio is mainly composed of middle-aged drivers and older vehicles. Premium per exposure and vehicle value are right-skewed, which is common in insurance portfolios because a small number of high-value or high-risk contracts can carry much higher premiums.

## 5.4 Claim Count and Incurred Loss

Claim variables need separate treatment because most records have zero claims, while incurred costs are highly right-skewed.

```{r}
claim_count_plot <- motor_analysis %>%
  count(claim_count_band, name = "records") %>%
  mutate(share = records / sum(records)) %>%
  ggplot(aes(x = claim_count_band, y = records)) +
  geom_col(fill = neutral_fill, width = 0.7) +
  geom_text(aes(label = fmt_pct(share)), vjust = -0.35, size = 3.2) +
  scale_y_continuous(
    labels = fmt_count,
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title = "Claim count per policy-year",
    x = "Total claims",
    y = "Policy-year records"
  ) +
  theme(plot.title = element_text(face = "bold"))

incurred_plot <- motor_analysis %>%
  filter(total_incurred > 0) %>%
  ggplot(aes(x = total_incurred)) +
  geom_histogram(bins = 50, fill = accent_fill, colour = "white", linewidth = 0.15) +
  scale_x_log10(labels = fmt_money) +
  scale_y_continuous(labels = fmt_count) +
  labs(
    title = "Positive incurred losses",
    subtitle = "Log scale is used because incurred losses have a long right tail.",
    x = "Total incurred, log scale",
    y = "Claim records"
  ) +
  theme(plot.title = element_text(face = "bold"))

claim_count_plot | incurred_plot
```

The no-claim share is `r fmt_pct(portfolio_metrics$no_claim_share)`, and the no-incurred share is `r fmt_pct(portfolio_metrics$no_incurred_share)`. This validates the use of zero-aware charts and concentration analysis in later sections.

## 5.5 Annualised Premium by Policy Type

Ridgeline density plots are useful for comparing the full premium distribution across several policy types, rather than comparing only means.

```{r}
set.seed(608)
sample_n <- min(nrow(motor_analysis), 100000)
motor_sample <- motor_analysis %>%
  filter(!is.na(premium_per_exposure), is.finite(premium_per_exposure)) %>%
  slice_sample(n = sample_n)

premium_cap <- quantile(motor_analysis$premium_per_exposure, 0.99, na.rm = TRUE)

ggplot(
  motor_sample %>%
    mutate(premium_plot = pmin(premium_per_exposure, premium_cap)),
  aes(x = premium_plot, y = policy_type, fill = policy_type)
) +
  geom_density_ridges(
    alpha = 0.65,
    scale = 0.9,
    rel_min_height = 0.01,
    colour = "white",
    linewidth = 0.25
  ) +
  scale_x_continuous(labels = fmt_money) +
  scale_fill_manual(values = c(
    "Third-party" = "#b7b7b7",
    "Third-party + glass" = "#9ab7c6",
    "Third-party + add-ons" = "#73b2c4",
    "Comprehensive with excess" = "#f2a65a",
    "Comprehensive without excess" = "#f27279"
  )) +
  labs(
    title = "Annualised premium distribution by policy type",
    subtitle = "The display is capped at the 99th percentile; summaries use the full data.",
    x = "Premium per exposure",
    y = NULL
  ) +
  guides(fill = "none") +
  theme(
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold")
  )
```

The premium distributions increase with coverage breadth, as expected. Comprehensive covers have wider and higher premium distributions than basic third-party products.

# 6. Exploratory Data Analysis

```{r}
summarise_segment <- function(data, group_var) {
  data %>%
    group_by({{ group_var }}) %>%
    summarise(
      records = n(),
      exposure = sum(total_exposure, na.rm = TRUE),
      premium = sum(total_premium, na.rm = TRUE),
      claims = sum(total_claims, na.rm = TRUE),
      incurred = sum(total_incurred, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      exposure_share = exposure / sum(exposure),
      loss_ratio = incurred / premium,
      claim_frequency = claims / exposure,
      claim_severity = if_else(claims > 0, incurred / claims, NA_real_),
      pure_premium = incurred / exposure,
      avg_premium_per_exposure = premium / exposure
    )
}

portfolio_lr <- portfolio_metrics$loss_ratio
```

## 6.1 Portfolio Trend by Year

The yearly view combines volume and profitability. Exposure is shown as bars, while loss ratio and claim frequency are shown as line charts because they are rates over time.

```{r}
year_summary <- summarise_segment(motor_analysis, year) %>%
  mutate(year_num = as.integer(as.character(year)))

p_year_exposure <- ggplot(year_summary, aes(x = year_num, y = exposure)) +
  geom_col(fill = base_fill, width = 0.65) +
  scale_x_continuous(breaks = year_summary$year_num) +
  scale_y_continuous(labels = fmt_count) +
  labs(
    title = "Exposure increased strongly from 2022 to 2024",
    x = NULL,
    y = "Exposure"
  ) +
  theme(plot.title = element_text(face = "bold", size = 10))

p_year_rates <- year_summary %>%
  select(year_num, loss_ratio, claim_frequency) %>%
  pivot_longer(-year_num, names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(
    metric,
    loss_ratio = "Loss ratio",
    claim_frequency = "Claim frequency"
  )) %>%
  ggplot(aes(x = year_num, y = value, colour = metric)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.8) +
  scale_x_continuous(breaks = year_summary$year_num) +
  scale_y_continuous(labels = fmt_rate) +
  scale_colour_manual(values = c("Loss ratio" = accent_fill, "Claim frequency" = dark_fill)) +
  labs(
    title = "Loss ratio worsened in 2024",
    x = NULL,
    y = "Rate",
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 10)
  )

p_year_exposure | p_year_rates
```

```{r}
year_display <- year_summary %>%
  transmute(
    Year = year,
    Exposure = fmt_count(exposure),
    Premium = fmt_money(premium),
    Claims = fmt_count(claims),
    Incurred = fmt_money(incurred),
    `Loss ratio` = fmt_pct(loss_ratio),
    `Claim frequency` = fmt_rate(claim_frequency)
  )

kable(year_display)
```

Exposure nearly triples across the period, from `r fmt_count(year_summary$exposure[year_summary$year == "2022"])` in 2022 to `r fmt_count(year_summary$exposure[year_summary$year == "2024"])` in 2024. The loss ratio also rises from `r fmt_pct(year_summary$loss_ratio[year_summary$year == "2022"])` to `r fmt_pct(year_summary$loss_ratio[year_summary$year == "2024"])`, suggesting that the expanding portfolio should be monitored for profitability drift.

## 6.2 Policy Type and Profitability

Policy type is one of the most important underwriting dimensions because it captures coverage breadth. The three metrics are plotted separately because loss ratio, claim frequency, and claim severity have very different scales.

```{r}
policy_summary <- summarise_segment(motor_analysis, policy_type)

policy_colours <- c(
  "Third-party" = "#b7b7b7",
  "Third-party + glass" = "#9ab7c6",
  "Third-party + add-ons" = "#73b2c4",
  "Comprehensive with excess" = "#f2a65a",
  "Comprehensive without excess" = "#f27279"
)

make_policy_bar <- function(data, metric, title_text, axis_labeller) {
  data %>%
    mutate(policy_type = fct_reorder(policy_type, .data[[metric]])) %>%
    ggplot(aes(x = .data[[metric]], y = policy_type, fill = policy_type)) +
    geom_col(width = 0.68) +
    geom_text(
      aes(label = axis_labeller(.data[[metric]])),
      hjust = -0.12,
      size = 3.1
    ) +
    scale_x_continuous(
      labels = axis_labeller,
      expand = expansion(mult = c(0, 0.18))
    ) +
    scale_fill_manual(values = policy_colours) +
    labs(
      title = title_text,
      x = NULL,
      y = NULL
    ) +
    guides(fill = "none") +
    theme(
      panel.grid.major.y = element_blank(),
      plot.title = element_text(face = "bold", size = 10)
    )
}

p_policy_lr <- make_policy_bar(
  policy_summary,
  "loss_ratio",
  "Loss ratio",
  fmt_pct
)

p_policy_freq <- make_policy_bar(
  policy_summary,
  "claim_frequency",
  "Claim frequency",
  fmt_rate
)

p_policy_sev <- make_policy_bar(
  policy_summary,
  "claim_severity",
  "Claim severity",
  fmt_money
)

p_policy_lr / p_policy_freq / p_policy_sev +
  plot_annotation(
    title = "Policy-type risk indicators by scale",
    subtitle = "Comprehensive without excess leads loss ratio and frequency;\nthird-party covers show higher severity when claims occur.",
    theme = theme(
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(size = 9)
    )
  )
```

```{r}
policy_display <- policy_summary %>%
  arrange(desc(loss_ratio)) %>%
  transmute(
    `Policy type` = policy_type,
    Exposure = fmt_count(exposure),
    `Exposure share` = fmt_pct(exposure_share),
    `Loss ratio` = fmt_pct(loss_ratio),
    `Claim frequency` = fmt_rate(claim_frequency),
    `Claim severity` = fmt_money(claim_severity)
  )

kable(policy_display)
```

`Comprehensive without excess` has the highest loss ratio at `r fmt_pct(max(policy_summary$loss_ratio, na.rm = TRUE))` and the highest claim frequency at `r fmt_rate(max(policy_summary$claim_frequency, na.rm = TRUE))`. This is consistent with the broader coverage design: more covered events can become payable claims. However, the result also signals that this product should be reviewed for deductible design, premium adequacy, and claim control.

## 6.3 Bonus Score and Claims Experience

The bonus score records past claims experience, so it should be directly related to future claim frequency if it is working as an underwriting signal.

```{r}
bonus_summary <- summarise_segment(motor_analysis, bonus_score)

bonus_long <- bonus_summary %>%
  select(bonus_score, loss_ratio, claim_frequency, avg_premium_per_exposure) %>%
  pivot_longer(-bonus_score, names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(
    metric,
    loss_ratio = "Loss ratio",
    claim_frequency = "Claim frequency",
    avg_premium_per_exposure = "Premium per exposure"
  ))

ggplot(bonus_long, aes(x = bonus_score, y = value, fill = bonus_score)) +
  geom_col(width = 0.7) +
  facet_wrap(~ metric, scales = "free_y", ncol = 1) +
  scale_fill_manual(values = c(
    "Poor history" = accent_fill,
    "Neutral history" = warning_fill,
    "Favourable history" = positive_fill
  )) +
  scale_y_continuous(labels = fmt_rate) +
  labs(
    title = "Bonus score separates risk, but premium relativities do not fully remove loss-ratio differences",
    x = NULL,
    y = NULL
  ) +
  guides(fill = "none") +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 20, hjust = 1),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )
```

Poor-history policies have a claim frequency of `r fmt_rate(bonus_summary$claim_frequency[bonus_summary$bonus_score == "Poor history"])`, compared with `r fmt_rate(bonus_summary$claim_frequency[bonus_summary$bonus_score == "Favourable history"])` for favourable-history policies. The pricing level is also much higher for poor-history policies, but their loss ratio remains above the favourable segment. This suggests that the bonus score is useful but may not fully capture residual risk.

## 6.4 Driver Age and Vehicle Age

```{r}
driver_age_summary <- summarise_segment(motor_analysis, driver_age_band) %>%
  filter(!is.na(driver_age_band)) %>%
  mutate(band = factor(as.character(driver_age_band), levels = levels(motor_analysis$driver_age_band)))

vehicle_age_summary <- summarise_segment(motor_analysis, vehicle_age_band) %>%
  filter(!is.na(vehicle_age_band)) %>%
  mutate(band = factor(as.character(vehicle_age_band), levels = levels(motor_analysis$vehicle_age_band)))

make_band_rate_plot <- function(summary_df, title_text) {
  summary_df %>%
    select(band, loss_ratio, claim_frequency, claim_severity) %>%
    pivot_longer(-band, names_to = "metric", values_to = "value") %>%
    mutate(metric = recode(
      metric,
      loss_ratio = "Loss ratio",
      claim_frequency = "Claim frequency",
      claim_severity = "Claim severity"
    )) %>%
    ggplot(aes(x = band, y = value, group = metric, colour = metric)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.2) +
    facet_wrap(~ metric, scales = "free_y", ncol = 1) +
    scale_colour_manual(values = c(
      "Loss ratio" = accent_fill,
      "Claim frequency" = dark_fill,
      "Claim severity" = warning_fill
    )) +
    scale_y_continuous(labels = fmt_rate) +
    labs(title = title_text, x = NULL, y = NULL, colour = NULL) +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 30, hjust = 1),
      plot.title = element_text(face = "bold", size = 10),
      strip.text = element_text(face = "bold")
    )
}

make_band_rate_plot(driver_age_summary, "Driver age bands") |
  make_band_rate_plot(vehicle_age_summary, "Vehicle age bands")
```

Claim frequency generally decreases with older driver and vehicle age bands, while severity is higher in the oldest driver and vehicle groups. This is a useful actuarial pattern: lower frequency does not always mean lower total risk if severity rises.

## 6.5 Geography and Coverage Interaction

```{r}
geo_policy_summary <- motor_analysis %>%
  mutate(geo_segment = paste(municipality_type, circulation_area, sep = " / ")) %>%
  group_by(policy_type, geo_segment) %>%
  summarise(
    exposure = sum(total_exposure, na.rm = TRUE),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(loss_ratio = incurred / premium)

ggplot(geo_policy_summary, aes(x = geo_segment, y = policy_type, fill = loss_ratio)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = fmt_pct(loss_ratio)), size = 3) +
  scale_fill_gradient2(
    low = "#73b2c4",
    mid = "#f5f5f5",
    high = "#d95f02",
    midpoint = portfolio_lr,
    labels = fmt_pct
  ) +
  labs(
    title = "Loss ratio by policy type and geographic driving segment",
    subtitle = "Colours are centred on the portfolio loss ratio.",
    x = NULL,
    y = NULL,
    fill = "Loss ratio"
  ) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold")
  )
```

The heatmap highlights that policy type differences are larger than geography differences, but some urban and coastal combinations show elevated loss ratios. This supports treating geography as a secondary segmentation layer rather than the sole driver of portfolio performance.

## 6.6 Vehicle Brand Concentration

The following bubble chart focuses on vehicle brands with at least 1,000 exposure years. This avoids over-interpreting very small brand segments.

```{r}
brand_summary <- summarise_segment(motor_analysis, vehicle_brand) %>%
  filter(exposure >= 1000) %>%
  slice_max(exposure, n = 15) %>%
  mutate(vehicle_brand = fct_reorder(vehicle_brand, loss_ratio))

ggplot(brand_summary, aes(
  x = loss_ratio,
  y = vehicle_brand,
  size = exposure,
  colour = claim_frequency
)) +
  geom_vline(xintercept = portfolio_lr, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.8) +
  scale_x_continuous(labels = fmt_pct) +
  scale_size_continuous(labels = fmt_count, range = c(3, 10)) +
  scale_colour_gradient(low = "#73b2c4", high = "#d95f02") +
  labs(
    title = "Large vehicle-brand segments differ in loss ratio and claim frequency",
    subtitle = "Dashed line marks the portfolio loss ratio.",
    x = "Loss ratio",
    y = NULL,
    size = "Exposure",
    colour = "Claim frequency"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold")
  )
```

Among the largest brand segments, several brands sit above the portfolio loss ratio. Because exposure is also high, these segments can materially affect the total result and deserve deeper underwriting review.

## 6.7 Premium Adequacy by Premium Decile

Premium-per-exposure deciles provide a simple calibration view. The first chart compares annualised premium with pure premium; the second chart checks whether the loss ratio remains stable across price bands.

```{r}
premium_decile_summary <- motor_analysis %>%
  filter(
    total_exposure > 0,
    total_premium > 0,
    !is.na(premium_per_exposure),
    is.finite(premium_per_exposure)
  ) %>%
  mutate(
    premium_decile = ntile(premium_per_exposure, 10),
    decile_label = factor(paste0("D", premium_decile), levels = paste0("D", 1:10))
  ) %>%
  group_by(premium_decile, decile_label) %>%
  summarise(
    records = n(),
    exposure = sum(total_exposure, na.rm = TRUE),
    premium = sum(total_premium, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    avg_premium_per_exposure = mean(premium_per_exposure, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    pure_premium = incurred / exposure,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure
  )

p_decile_money <- premium_decile_summary %>%
  select(decile_label, avg_premium_per_exposure, pure_premium) %>%
  pivot_longer(-decile_label, names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(
    metric,
    avg_premium_per_exposure = "Premium per exposure",
    pure_premium = "Pure premium"
  )) %>%
  ggplot(aes(x = decile_label, y = value, colour = metric, group = metric)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_y_continuous(labels = fmt_money) +
  scale_colour_manual(values = c("Premium per exposure" = dark_fill, "Pure premium" = accent_fill)) +
  labs(
    title = "Premium and pure premium increase across pricing deciles",
    x = NULL,
    y = "Amount per exposure",
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 10)
  )

p_decile_lr <- ggplot(premium_decile_summary, aes(x = decile_label, y = loss_ratio)) +
  geom_col(fill = base_fill, width = 0.7) +
  geom_hline(yintercept = portfolio_lr, linetype = "dashed", colour = "grey40") +
  scale_y_continuous(labels = fmt_pct) +
  labs(
    title = "Loss ratio is not perfectly flat across deciles",
    x = "Premium per exposure decile",
    y = "Loss ratio"
  ) +
  theme(plot.title = element_text(face = "bold", size = 10))

p_decile_money / p_decile_lr
```

Higher premium deciles do carry higher claim frequency and pure premium, which indicates that pricing is directionally risk-sensitive. However, the loss ratio is not monotonic across deciles. This suggests that premium level captures part of the risk signal, but some segments may still be over- or under-priced.

## 6.8 Volatility and Loss Concentration

Insurance portfolios can be volatile because large losses are rare but expensive. A Pareto-style cumulative curve is appropriate because it shows how much incurred cost is explained by the highest-loss policy-year records.

```{r}
incurred_sorted <- motor_analysis %>%
  arrange(desc(total_incurred)) %>%
  mutate(
    rank = row_number(),
    policy_share = rank / n(),
    incurred_share = cumsum(total_incurred) / sum(total_incurred, na.rm = TRUE)
  )

top_share_tbl <- tibble(policy_share = c(0.001, 0.005, 0.01, 0.05, 0.10)) %>%
  mutate(
    records = ceiling(policy_share * nrow(motor_analysis)),
    incurred_share = map_dbl(
      records,
      ~ sum(incurred_sorted$total_incurred[seq_len(.x)], na.rm = TRUE) /
        sum(motor_analysis$total_incurred, na.rm = TRUE)
    )
  )

pareto_data <- incurred_sorted %>%
  filter(rank == 1 | rank %% 500 == 0 | rank == n())

ggplot(pareto_data, aes(x = policy_share, y = incurred_share)) +
  geom_abline(linetype = "dashed", colour = "grey70") +
  geom_line(colour = accent_fill, linewidth = 1) +
  geom_point(
    data = top_share_tbl %>% filter(policy_share %in% c(0.01, 0.05)),
    aes(x = policy_share, y = incurred_share),
    inherit.aes = FALSE,
    colour = dark_fill,
    size = 2.6
  ) +
  geom_text(
    data = top_share_tbl %>% filter(policy_share %in% c(0.01, 0.05)),
    aes(
      x = policy_share,
      y = incurred_share,
      label = paste0(fmt_pct(policy_share), " of records\n", fmt_pct(incurred_share), " of incurred")
    ),
    inherit.aes = FALSE,
    hjust = -0.05,
    vjust = 1,
    size = 3
  ) +
  scale_x_continuous(labels = fmt_pct, limits = c(0, 1)) +
  scale_y_continuous(labels = fmt_pct, limits = c(0, 1)) +
  labs(
    title = "Incurred losses are highly concentrated",
    subtitle = "Policy-year records are sorted from highest to lowest incurred loss.",
    x = "Cumulative share of policy-year records",
    y = "Cumulative share of incurred losses"
  ) +
  theme(plot.title = element_text(face = "bold"))
```

```{r}
top_share_display <- top_share_tbl %>%
  transmute(
    `Top share of records` = fmt_pct(policy_share),
    `Number of records` = fmt_count(records),
    `Share of incurred losses` = fmt_pct(incurred_share)
  )

kable(top_share_display)
```

The top `r fmt_pct(top_share_tbl$policy_share[top_share_tbl$policy_share == 0.01])` of policy-year records account for `r fmt_pct(top_share_tbl$incurred_share[top_share_tbl$policy_share == 0.01])` of incurred losses. This is the clearest evidence of portfolio volatility and explains why severity monitoring is as important as average loss-ratio monitoring.

## 6.9 Statistical Cross-Checks

The visual patterns above are supported by simple statistical checks. Given the very large sample size, p-values should not be interpreted alone; the visual effect size and business relevance remain more important.

```{r}
test_tbl <- tribble(
  ~Question, ~Test, ~P_value,
  "Is claim occurrence independent of policy type?",
  "Chi-squared test",
  chisq.test(table(motor_analysis$claim_indicator, motor_analysis$policy_type))$p.value,
  "Is claim occurrence independent of bonus score?",
  "Chi-squared test",
  chisq.test(table(motor_analysis$claim_indicator, motor_analysis$bonus_score))$p.value,
  "Is claim occurrence independent of circulation area?",
  "Chi-squared test",
  chisq.test(table(motor_analysis$claim_indicator, motor_analysis$circulation_area))$p.value,
  "Do capped policy-year loss ratios differ by policy type?",
  "Kruskal-Wallis test",
  kruskal.test(
    pmin(loss_ratio, quantile(loss_ratio, 0.99, na.rm = TRUE)) ~ policy_type,
    data = motor_analysis
  )$p.value
)

test_tbl %>%
  mutate(P_value = format.pval(P_value, digits = 3, eps = 0.001)) %>%
  kable()
```

The tests reject independence or equality for the selected relationships at conventional thresholds. This supports the observed visual differences, especially for policy type and bonus score. However, because the portfolio is large, these tests mainly confirm that the patterns are unlikely to be random; business action should still be guided by the magnitude of the differences.

## 6.10 Interactive Visual Analytics

### 6.10.1 Searchable Segment Performance Table

The DT table allows readers to sort and filter policy, year, bonus-score, and status segments by exposure, loss ratio, claim frequency, or severity.

```{r}
interactive_segment_table <- bind_rows(
  year_summary %>%
    transmute(
      `Segment group` = "Year",
      Segment = as.character(year),
      records, exposure, premium, claims, incurred,
      loss_ratio, claim_frequency, claim_severity
    ),
  policy_summary %>%
    transmute(
      `Segment group` = "Policy type",
      Segment = as.character(policy_type),
      records, exposure, premium, claims, incurred,
      loss_ratio, claim_frequency, claim_severity
    ),
  bonus_summary %>%
    transmute(
      `Segment group` = "Bonus score",
      Segment = as.character(bonus_score),
      records, exposure, premium, claims, incurred,
      loss_ratio, claim_frequency, claim_severity
    ),
  summarise_segment(motor_analysis, policy_status) %>%
    transmute(
      `Segment group` = "Policy status",
      Segment = as.character(policy_status),
      records, exposure, premium, claims, incurred,
      loss_ratio, claim_frequency, claim_severity
    )
) %>%
  rename(
    Records = records,
    Exposure = exposure,
    Premium = premium,
    Claims = claims,
    Incurred = incurred,
    `Loss ratio` = loss_ratio,
    `Claim frequency` = claim_frequency,
    `Claim severity` = claim_severity
  )

DT::datatable(
  interactive_segment_table,
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE
  )
) %>%
  DT::formatRound(
    columns = c("Records", "Exposure", "Premium", "Claims", "Incurred", "Claim severity"),
    digits = 1
  ) %>%
  DT::formatRound(columns = "Claim frequency", digits = 3) %>%
  DT::formatPercentage(columns = "Loss ratio", digits = 1)
```

### 6.10.2 Interactive Geographic Heatmap

The static heatmap already identifies high-loss-ratio cells. The interactive version adds hover details, so readers can check whether a cell is material by exposure and premium volume.

```{r}
geo_policy_interactive <- geo_policy_summary %>%
  mutate(
    hover_text = paste0(
      "<b>", policy_type, " | ", geo_segment, "</b>",
      "<br>Exposure: ", fmt_count(exposure),
      "<br>Premium: ", fmt_money(premium),
      "<br>Incurred: ", fmt_money(incurred),
      "<br>Claims: ", fmt_count(claims),
      "<br>Loss ratio: ", fmt_pct(loss_ratio)
    )
  )

geo_heatmap_plot <- ggplot(
  geo_policy_interactive,
  aes(x = geo_segment, y = policy_type, fill = loss_ratio, text = hover_text)
) +
  geom_tile(colour = "white", linewidth = 0.5) +
  scale_fill_gradient2(
    low = "#73b2c4",
    mid = "#f5f5f5",
    high = "#d95f02",
    midpoint = portfolio_lr,
    labels = fmt_pct
  ) +
  labs(
    title = "Interactive loss-ratio heatmap",
    x = NULL,
    y = NULL,
    fill = "Loss ratio"
  ) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold")
  )

ggplotly(geo_heatmap_plot, tooltip = "text") %>%
  layout(legend = list(orientation = "h", x = 0, y = -0.35))
```

### 6.10.3 Interactive Vehicle-Brand Bubble Chart

The reader can hover over a brand to see the exact exposure, loss ratio, claim frequency, and severity.

```{r}
brand_interactive <- brand_summary %>%
  mutate(
    tooltip = paste0(
      vehicle_brand,
      "\nExposure: ", fmt_count(exposure),
      "\nLoss ratio: ", fmt_pct(loss_ratio),
      "\nClaim frequency: ", fmt_rate(claim_frequency),
      "\nClaim severity: ", fmt_money(claim_severity)
    )
  )

brand_giraph_plot <- ggplot(
  brand_interactive,
  aes(x = loss_ratio, y = vehicle_brand, size = exposure, colour = claim_frequency)
) +
  geom_vline(xintercept = portfolio_lr, linetype = "dashed", colour = "grey40") +
  geom_point_interactive(
    aes(tooltip = tooltip, data_id = vehicle_brand),
    alpha = 0.85
  ) +
  scale_x_continuous(labels = fmt_pct) +
  scale_size_continuous(labels = fmt_count, range = c(3, 10)) +
  scale_colour_gradient(low = "#73b2c4", high = "#d95f02") +
  labs(
    title = "Interactive brand concentration view",
    subtitle = "Hover over a point to inspect segment metrics.",
    x = "Loss ratio",
    y = NULL,
    size = "Exposure",
    colour = "Claim frequency"
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold")
  )

girafe(
  ggobj = brand_giraph_plot,
  width_svg = 8,
  height_svg = 5,
  options = list(
    opts_hover(css = "stroke:#244c5a;stroke-width:2px;cursor:pointer;"),
    opts_tooltip(css = "background-color:white;color:#244c5a;padding:6px;border:1px solid #c7c8cc;border-radius:4px;")
  )
)
```

### 6.10.4 Interactive Loss-Concentration Curve

The Plotly Pareto curve makes it easier to inspect how quickly incurred losses accumulate as the highest-loss records are added.

```{r}
pareto_hover <- pareto_data %>%
  mutate(
    hover_text = paste0(
      "Rank: ", fmt_count(rank),
      "<br>Cumulative records: ", fmt_pct(policy_share),
      "<br>Cumulative incurred: ", fmt_pct(incurred_share)
    )
  )

top_share_hover <- top_share_tbl %>%
  mutate(
    hover_text = paste0(
      "Top records: ", fmt_pct(policy_share),
      "<br>Number of records: ", fmt_count(records),
      "<br>Share of incurred: ", fmt_pct(incurred_share)
    )
  )

plot_ly() %>%
  add_lines(
    data = pareto_hover,
    x = ~ policy_share,
    y = ~ incurred_share,
    text = ~ hover_text,
    hoverinfo = "text",
    line = list(color = accent_fill, width = 3),
    name = "Cumulative incurred"
  ) %>%
  add_lines(
    x = c(0, 1),
    y = c(0, 1),
    line = list(color = "gray70", dash = "dash"),
    hoverinfo = "skip",
    showlegend = FALSE
  ) %>%
  add_markers(
    data = top_share_hover,
    x = ~ policy_share,
    y = ~ incurred_share,
    text = ~ hover_text,
    hoverinfo = "text",
    marker = list(color = dark_fill, size = 8),
    name = "Selected thresholds"
  ) %>%
  layout(
    title = "Interactive incurred-loss concentration curve",
    xaxis = list(title = "Cumulative share of policy-year records", tickformat = ".0%"),
    yaxis = list(title = "Cumulative share of incurred losses", tickformat = ".0%"),
    legend = list(orientation = "h", x = 0, y = -0.2)
  )
```

# 7. Summary

This visual analysis leads to the following evidence-based findings:

- The portfolio is large and growing: exposure increased from 2022 to 2024, while the loss ratio rose to `r fmt_pct(year_summary$loss_ratio[year_summary$year == "2024"])` in 2024.
- The overall loss ratio is `r fmt_pct(portfolio_metrics$loss_ratio)`, but profitability is uneven across segments.
- `Comprehensive without excess` is the most concerning policy type, with the highest claim frequency and the highest loss ratio.
- Bonus score is a meaningful underwriting signal: poor-history policies have materially higher claim frequency than favourable-history policies.
- Age-band analysis shows why both frequency and severity must be monitored. Some older groups claim less often, but when they do claim, average incurred cost can be higher.
- Geography is a secondary but useful segmentation layer. Policy type explains larger differences, while some urban and coastal combinations show elevated loss ratios.
- Premium level is directionally risk-sensitive, but loss ratio is not perfectly stable across premium deciles, suggesting room for pricing refinement.
- Losses are highly concentrated: the top `r fmt_pct(top_share_tbl$policy_share[top_share_tbl$policy_share == 0.01])` of records explain `r fmt_pct(top_share_tbl$incurred_share[top_share_tbl$policy_share == 0.01])` of incurred losses. This is the main source of portfolio volatility.
- Interactive visualisations are most useful at the segment level for this dataset. The added DT, Plotly, and ggiraph outputs help readers inspect segment details without plotting every policy-year record.

From an underwriting and portfolio-management perspective, the most practical next steps are to investigate comprehensive-without-excess policies, review bonus-score relativities, monitor 2024 profitability drift, and build a severity-monitoring view for large incurred losses.

# 
